### 2022-11-13 article.Rmd
Description of a reuse networks, and some layout tweaks

---
title: "Two years of CiTO annotations"
subtitle: "Type: Correspondance"
author:
- Egon Willighagen [1,*,orcid:0000-0001-7542-0286]
link-citations: yes
geometry: margin=2.5cm
output:
  html_document:
    pandoc_args:
    - --lua-filter=filters/extract-cito.lua
    - --citeproc
    - --lua-filter=filters/insert-cito-in-ref.lua
  pdf_document:
    pandoc_args:
    - --lua-filter=filters/extract-cito.lua
    - --citeproc
    - --lua-filter=filters/insert-cito-in-ref.lua
  word_document:
    pandoc_args:
    - --lua-filter=filters/extract-cito.lua
    - --citeproc
    - --lua-filter=filters/insert-cito-in-ref.lua
# RStudio: cito-bibliography: bibliography.bib
# Other: bibliography: bibliography.bib
cito-bibliography: bibliography.bib
csl: journal-of-cheminformatics.csl
...

* Correspondance:
egon.willighagen@maastrichtuniversity.nl

# Abstract

Citations are an essential aspect of research communication and have become the basis of many
metrics in the academic world. Some see citation counts as a mark of scientific impact or
even quality, but in reality the reasons for citing other work are manifold which makes
the interpretation more complicated than a single citation count can reflect.
Two years ago, the *Journal of Cheminformatics* proposed the CiTO Pilot for the adoption of
a practice of annotating citations with their citation intentions.
Basically, when you cite a journal article or dataset (or any other soruce), you also explain
why specifically you cite that source. Particularly,
the agreement and disagreement and reuse of methods and data are of interest.
This article explores what happened after the launch of the pilot. We summarize how
authors in the *Journal of Cheminformatics* used the pilot, shows citation annotations
are distributed with Wikidata, visualized with Scholia, discusses adoption outside BioMedCentral,
and ponders about what needs to happen next.

# Main text

Communiting new research findings is primarily done by written texts in the form of scholarly
articles, books, and book chapters. To not having to repeat past research by themselves or others,
authors cite research relevant [@agreesWith:citesAsAuthority:Shotton_2010]. However, the reasons why
authors cite literature vary, which complicates how we use citations [@agreesWith:citesAsAuthority:Nicholson_2021]. 
Typing citations is therefore of interest: it allows us to navigate ltierature more easily: it points
us to essential research methods, data, and can warn us of research that cannot be reproduced,
or others disagree with [@Nicholson_2021].

With the use of citations increasingly being picked up to help researchers with tools like
scite.ai [@citesAsAuthority:Nicholson_2021] en Connected Papers [@citesAsAuthority:ConnectedPapers], having typed
citations will help us explore literature. Therefore, the *Journal of Cheminformatics*
started a pilot for the adoption of annotating citations with their annotations [@discusses:Willighagen_2020].

# The Citation Typing Ontology Pilot

The pilot consists of a couple of components and the editorial explains some of them [@Willighagen_2020].
The Citation Typing Ontology was selected to express the intention [@usesMethodIn:Shotton_2010],
the intention is expressed a compact identifier wrapped in square brackets, also called a
safe CURIE, standard proposed by the W3C in 2010 [@usesMethodIn:McCarron2010,@usesMethodIn:Wimalaratne_2018].
The *cito* prefix is registered in BioRegistry [@usesDataFrom:Hoyt2022].
The *bibnotes* concept of the Springer Nature publishing platform was used as carrier.
Authors are guided by a landing page consisting of a BioMedCentral Collection at
[https://www.biomedcentral.com/collections/cito](https://www.biomedcentral.com/collections/cito)
and author guidelines explaining to authors
how they can add the annotations with their favoriate editor at
[https://jcheminform.github.io/jcheminform-author-guidelines/cito](https://jcheminform.github.io/jcheminform-author-guidelines/cito).

Because the CiTO ontology has many terms for many different citation intentions,
we made a selection of CiTO terms authors could use [@CiTOCollection]:
**[cito:citesAsDatasource]** to indicate
to a source the provides data to back up a claim, **[cito:usesDataFrom]** to indicate that the authors
reused data, **[cito:usesMethodIn]** when a method or protocol explain in that source is used,
and a few more general intentions like **[cito:discusses]**, **[cito:extends]**, **[cito:agreesWith]**,
and **[cito:disagreesWith]**.
The journal itself would adopt the following CiTO annotations: **[cito:retracts]**,
**[cito:repliesTo]**, and **[updates]**. Fortunately, it has not been used yet, but
the first would be used if an article with be retracted from the journal. The second
would be used when an Letter to the Editor replies to an earlier published article,
and **[cito:updates]** when an Correction was published.

# Wikidata and Scholia

To track the uptake but also to demonstrate the impact, we extended Scholia to visualize
citation intention data. Scholia is a graphical interface around the data stored in
Wikidata [@usesMethodIn:Nielsen_2017] and includes citations from OpenCitations [@Peroni_2020]
and PubMed. Wikidata allows adding qualifier to statements which allowed us to define
a data model for citations annoted with CiTO intention; the Wikidata property P3712
has been used for this, labeled *objective of project or action* (see Figure 1).

![Screenshot of the citation statements for an article where the
*objective of project or action* qualifier is used to annotate the citation with their
CiTO intentions.](images/datamodel.png)

This data in Wikidata can then be accessed via Scholia and Scholia tells use some
overall statisticss of the number of annotations, which we reported on about a year
ago too [@citesAsDataSource:Willighagen_2021]. Since last year and written down
on August 25, the number of annotations and the number of annotated citations
have almost doubled (from 377 to 603 and from 304 to 494, respectively). The first
number is higher because one citation can have more than one citation intention.
To continue, the current number of citations are citing 387 articles in 141 different
scholarly journals, and they are found in 98 articles in 48 different journals
(see Figure 2) [@citesAsDataSource:scholiaCiTOStatistics].

It must be noted that the *Journal of Cheminformatics* is only one possible source,
but is still the only journal that uses CiTO annotation explicitly in the articles itself.
And with 335 annotated citations in 32 articles it also is the major source of
CiTO annotations in Wikidata at the time of writing.
However, CiTO intention annotations in Wikidata can come from other sources too
and be added both manually and automatically using the tools around Wikidata.
When all annotation is combined, Scholia shows us that **[cito:citesAsAuthority]**
is the most used intention, with 226 annotated citations (out of 603) in 38 articles.
**[cito:usesMethodIn]** follows with 102 annotated citations.

# Adoption by *Journal of Cheminformatics* authors

In the first two years after the start of the Pilot, including the seminal editorial [@extends:Willighagen_2020],
the *Journal of Cheminformatics* has published fifteen articles with explicit CiTO annotation:
three Editorials, four Research articles, two Database, and one of each of Data Note, Software,
Letter to the Editor, Letter Response, Educational, and Methodology. Ten were published in
the first year (Table 1) and five in the second year (Table 2). Each article annotated
one or more citations with CiTO intentions, and several articles annotated every citation,
far exceeding the original anticipation.

![Screenshot of the Scholia Citation Typing Ontology page showing the
daily statistics.](images/scholiaStats.png)

Also exceeding expectations is the diversity of the chosen CiTO intention types. The original
guidance focused on **[cito:citesAsDataSource]**, **[cito:usesDataFrom]** (the first is used
to cite an article with data, the second when you reused data and citate the article where it
comes from), **[cito:usesMethodIn]**, **[cito:citesAsAuthority]**, **[cito:discusses]**,
**[cito:extends]**, **[cito:agreesWith]**, and **[cito:disagreesWith]**. Not only have
all of these been used, authors also used **[cito:citesForInformation]**,
**[cito:citesAsPotentialSolution]**, **[cito:citesAsRelated]**, **[cito:documents]**,
and **[cito:obtainsBackgroundFrom]**.

|**Type**  |**Article** |**Intentions** | **%CiTO** |
----- |------------ |------------ |---
|Editorial|Adoption of the Citation Typing Ontology by the *Journal of Cheminformatics* [@citesForInformation:Willighagen_2020]|cito:citesAsAuthority, cito:citesForInformation, cito:usesMethodIn|42%|
|Methodology|Predicting target profiles with confidence as a service using docking scores [@citesForInformation:Ahmed2020]|cito:agreesWith, cito:extends, cito:citesAsAuthority, cito:citesAsDataSource, cito:usesMethodIn|47%|
|Research article|Too sweet: cheminformatics for deglycosylation in natural products [@citesForInformation:Schaub2020]|cito:citeAsAuthority, cito:citesForInformation, cito:usesDataFrom, cito:usesMethodIn||
|Educational|A ligand-based computational drug repurposing pipeline using KNIME and Programmatic Data Access: case studies for rare diseases and COVID-19 [@citesForInformation:Tuerkova2020]|cito:agreesWith, cito:citesAsAuthority, cito:citesAsDataSource, cito:discusses, cito:usesDataFrom, cito:usesMethodIn|98%|
|Database|COCONUT online: Collection of Open Natural Products database [@citesForInformation:Sorokina2021]|cito:citesForInformation, cito:usesDataFrom, cito:usesMethodIn|41%|
|Editorial|What is the role of cheminformatics in a pandemic? [@citesForInformation:Guha2021]|cito:agreesWith, cito:citesAsAuthority, cito:citesAsDataSource, cito:citesForInformation, cito:citesAsPotentialSolution|91%|
|Research article|Empowering large chemical knowledge bases for exposomics: PubChemLite meets MetFrag [@citesForInformation:Schymanski2021]|cito:citesAsAuthority, cito:citesAsDataSource, cito:citesAsMetadataDocument, cito:discusses, cito:extends, cito:usesDataFrom, cito:usesMethodIn|100%|
|Software|IDSM ChemWebRDF: SPARQLing small-molecule datasets [@citesForInformation:Galgonek2021]|cito:citesAsAuthority, cito:citesAsRelated, cito:usesDataFrom, cito:usesMethodIn|97%|
|Letter to the Editor|FAIR chemical structures in the Journal of Cheminformatics [@citesForInformation:SchymanskiBolton2021]|cito:citesAsAuthority, cito:documents|100%|
|Letter Response|Reply to “FAIR chemical structure in the Journal of Cheminformatics” [@citesForInformation:Guha2021b]|cito:agreesWith, cito:citesAsAuthority, cito:citesAsDataSource, cito:obtainsBackgroundFrom, cito:repliesTo| 100%|

Table: *Journal of Cheminformatics* articles in the first year of the pilot.

|**Type**  |**Article** |**Intentions** | **%CiTO** |
----- |------------ |------------ |---
|Research article|DECIMER 1.0: deep learning for chemical image recognition using transformers [@citesForInformation:Rajan2021]|cito:agreesWith, cito:citesAsAuthority, cito:citesAsDataSource, cito:extends, cito:usesMethodIn|66% (*)|
|Research article|*DrugEx* v2: de novo design of drug molecules by Pareto-based multi-objective reinforcement learning in polypharmacology[@citesForInformation:Liu2021]|cito:extends, cito:citesAsAuthority, cito:usesMethodIn|100%|
|Database|PSnpBind: a database of mutated binding site protein–ligand complexes constructed using a multithreaded virtual screening workflow [@citesForInformation:Ammar2022]|cito:citesAsDataSource, cito:usesDataFrom, cito:usesMethodIn|23%|
|Editorial|Diversifying cheminformatics [@citesForInformation:Zdrazil2022]|cito:citesAsAuthority, cito:citesAsEvidence, cito:citesForInformation, cito:citesAsRecommendedReading, cito:citesAsSourceDocument, cito:containsAssertionFrom|100%|
|Data Note|DECIMER - hand-drawn molecule images dataset [@citesForInformation:Brinkhaus2022]|cito:agreesWith, cito:cites, cito:citesAsAuthority, cito:extends, cito:usesDataFrom, cito:usesMethodIn|41% (*)|

Table: *Journal of Cheminformatics* articles in the second year of the pilot. The * indicates that
the percentage is based on CiTO intends different from cito:cites.

# Technological innovation

To make life easier for authors, and following a Twitter discussion in Spring 2021, a Markdown template was developed with native CiTO support: https://jcheminform.github.io/jcheminform-author-guidelines/cito-guidelines/markdown.html. Here, the author merely indicates the CiTO type when they cite the article. This is using a method introduced by Krewinkel et al. [@usesMethodIn:Krewinkel2017].
The manuscript can then be converted to a Microsoft Word file with Pandoc (https://pandoc.org/) for submission to the journal.

The *Journal of Cheminformatics* template is available from the journal's GitHub organization, and authors and editors should feel free to adapt it to the needs of other journals. The BioHackrXiv ([https://biohackrxiv.org/](https://biohackrxiv.org/))
preprint server also support CiTO annotations [@citesAsDataSource:Prins2021] and this template can be found at
[https://github.com/biohackrxiv/publication-template](https://github.com/biohackrxiv/publication-template).

# Annotated Citation Networks

We already use citation networks in finding relevant literature, for example based on co-citation patterns.
Such analyses become stronger when we know more why articles cited. Similarly, an article that cites an
article because it uses a method in second article and that method extends a method in a third articles,
then the first article indirectly uses the method in that third article, even if that third article is
not directly cited. This is reflecting in citation habits: authors always decide whether to cite all
articles about a method, only the most recent, or only the oldest (or something else). After all, journals
frequently have rules about the maximum length of reference sections. Moreover, some methods are
so well established, we are not expected to cite that work at all.

![Citation network consisting only of citations annotated with **[cito:usesMethodIn]**.
Articles with explicit CiTO annotation are shown in red. The Wikidata SPARQL query
used to create this diagram is available from
[https://gist.github.com/egonw/1c8bc99373a24075838ee19976c74856](https://gist.github.com/egonw/1c8bc99373a24075838ee19976c74856).](images/usesMethodInNetwork.png)

The general availability of open citation allows us to recover such more complex reuse scenarios
using the citation networks. Moreover, when the citations are annotated, we can zoom in on reuse
networks. Figure 3 shows a method reuse network for articles with explicit annotation in red. The
network shows a few article that use methods in two cited articles, e.g. *Kohulan Rajan, 2021*
and *Henning Otto Brinkhaus, 2022* [@Rajan2021,@Brinkhaus2022]. Of course, if we do not limit
the network to a subset of articles (here, *Journal of Cheminformatics* articles with CiTO
annotation), the network becomes more interesting, but also much more complex. Network analysis
approaches can then be used instead of network visualization.

When we combined the reuse of methods and data, we can for journals summarize which articles
are most reused. Analyses like this becomes a simple query that can be routinely performed for
any journal. Figure 4 shows a tabular summary of articles of which methods or data is reused.
Like citation counts, this data depends totally on explicit citation data. However, these
citation counts are based on actual reuse and not also on the number of citations as authority.

![Screenshot of the "Most reused articles" section of the Scholia page for the
*Journal of Cheminformatics*. Based on CiTO annotation available in Wikdiata, the ten
most reused articles are shown.](images/mostUsedInJCheminf.png)

# Discussion

Because many publishing platforms currently do not support display of citation level intention annotation,
the default model only provides the annotation in the bibliography. The choice was made to use the `bibnotes`
idea which allows giving additional notes to a reference in the bibliography. The CiTO Pilot uses
the *safe CURIE* standard, but the typesetting in the *Journal of Cheminformatics* has shown to be inconsistent,
even within single articles, suggesting additional training or guidance may be needed.

This links to a limitation to the current use of CiTO annotation: the citation-level annotation may be
supported by some authoring systems (Markdown), even then the depiction may not be supported. If we convert
a Markdown file to a Microsoft Word file, the annotation is kept in the bibliography. However, when
writing a manuscript in Microsoft Word, LibreOffice, or Google Docs, combined with a reference manager
like Zotero, the CiTO annotation cannot be stored as part of the manuscript easily. The problem here is
that CiTO annotation in the reference manager are no longer linked to when they are cited. The workaround
is to add the CiTO annotation after completion of the Word document, directly to the bibliography.

LaTeX users may find them in a situation between that of Markdown users and reference manager users:
only if a manuscript-level Bib(La)TeX file is used the CiTO annotation can be added as notes to the
BiBTeX file. This is explained in this guidance document:
[https://jcheminform.github.io/jcheminform-author-guidelines/cito-guidelines/latex.html](https://jcheminform.github.io/jcheminform-author-guidelines/cito-guidelines/latex.html).

# What is next?

With fourteen articles published in the CiTO Collection, the adoption is not as high as one would have hoped.
Nevertheless, the pilot is triggering interest with both authors and publishing platforms. The
*Journal of Cheminformatics* has already published a few more articles with CiTO annotation and
a search on the preprint servers ResearchSquare and ChemrXiv show a few more manuscripts. The support
by BioHackrXiv is a nice example of adoption beyond BioMedCentral.

The Scholia
use case shows how the information can be used downstream. Further uptake of this idea of typed citations
depends on the combined willingness of journal editors, authors, publishers and indexing services alike.
The rise of services like scite.ai shows that the research community is ready this kind of information.
Logical steps include support of distributing citation typing annotation via platforms like CrossRef
or EuropePMC.

Because all innovation requires a critical amount of adoption to be accepted, additional sources of
CiTO annotation will be welcome. For example, providing CiTO annotation via nanopublications would
allow collections of annotations to be shared. Then, when archived on Zenodo and therefore citatble,
these annotations can be included in analysis as a trusted or at least citable source.

### Availability of data and materials

CiTO annotation in the *Journal of Cheminformatics* are available from the journal's articles. CiTO
annotation data in Scholia is available from [Wikidata](https://wikidata.org/). Markdown templates that
support CiTO are available from
[https://github.com/jcheminform/markdown-jcheminf](https://github.com/jcheminform/markdown-jcheminf) and
[https://github.com/biohackrxiv/publication-template](https://github.com/biohackrxiv/publication-template).

### Competing interests

No competing interests.

### Funding

Part of this work was supported by ELIXIR, the research infrastructure for life-science data.

### Authors' contributions

Not applicable.

### Acknowledgments

This work would not be possible without the support from Springer Nature and
Matthew Smyllie in particular and the editors of the *Journal of Cheminformatics*,
Rajarshi Guha, Nina Jealizakova, and Barbara Zdrazil. Carlin MacKenzie is thanked for
integration the sections on CiTO use into the main Scholia journal aspects [@citesAsEvidence:0af60cf7732bc8664d828cbe51f233aa63201df9].
Huge thanks goes to Albert Krewinkel for developing the Markdown/Pandoc integration.
Finally, thanks to all authors for including CiTO annotation in their articles.

# References